---
title: 'R4 Assignment: Optimization & Prescriptive Analytics'
author: "Daniel Saul & Michael Morgan"
date: "11/18/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Prepare Data & Files
```{r}
# Download required packages
library(tidyverse)
library(knitr)
library(lubridate)
library(tinytex)
library(lpSolveAPI)
library(TSP)
library(ggmap)
library(mapproj)
library(measurements)
library(googleway)
```

# A. Small Farmer and Crop Production [Blending]
```{r}
# Set up Optimization Equation and Solve
lp <- make.lp(0,2) #Starting with 0 constraints and 2 variables
set.objfn(lp, c(35,25))
add.constraint(lp, c(2, 3), "<=", 60)
add.constraint(lp, c(5, 2), "<=", 80)
add.constraint(lp, c(3, 2), "<=", 60)
add.constraint(lp, c(1, 2), "<=", 40)

#Set objective Direction 
invisible(lp.control(lp, sense = 'max'))
print(lp)

#Solve
solve(lp)

#Show Results
#Get Max Profit:
p <- get.objective(lp)
paste0('Max Profit: $',p[1])

#Get variables from maximizing
v <- get.variables(lp)

#Total units by Crop for maximum profit
paste0('Bushels of Crop A Produced: ', v[1])
paste0('Bushels of Crop B Produced: ', v[2])
paste0('Total Crops Produced: ', v[1] + v[2])

#Allocation of Resources to Crops
paste0('Total Land Units: ', 2*v[1] + 3*v[2])
paste0('Total Water Units: ', 5*v[1] + 2*v[2])
paste0('Total Fertilizer Units: ', 3*v[1] + 2*v[2])
paste0('Total Labor Units: ', 1*v[1] + 2*v[2])

#Shadow Prices
#Solve for additional resources needed
d <- get.dual.solution(lp)
d

#Show interpretations
paste0('Increasing one unit of land will increase max profit by: $', d[2])
paste0('Increasing one unit of water will increase max profit by: $', d[3])
paste0('Increasing one unit of fertilizer will make no change to max profit: ', d[4])
paste0('Increasing one unit of labor will make no change to max profit: ', d[5])
```

__How much of each crop should be produced and what's the maximum profit obtained with this distribution?__

10.91 Bushels of Crop A and 12.73 Bushels of Crop B should be produced in order to obtain the maximum profit of $700.

__What allocation of resources to crops will maximize the farmer's profit?__

The allocation of resources to crops for maximizing profit is as follows:

_Total Land Units:_ 60

_Total Water Units:_ 80

_Total Fertilizer Units:_ 58.1818181818182

_Total Labor Units:_ 36.3636363636364

__Use shadow price analysis to determine what additional resources the farmer should seek?__

The farmer should seek additional units of land and water, as these both will marginally increase max profit by $5.

# B. Pizza Chain and Orders [Traveling Salesman]
```{r}
k <- "AIzaSyBAScaGsIEmn4c0Oeg026ZhHcU_VXVh50M"

# TESTING STORE 1
store1 <- c("700 E. Broad, Athens, GA", "15 Ruthwood Lane, Athens, GA", "100 Kentucky Circle, Athens, GA", "558 West Broad St, Athens, GA", "125 Greek Park Circle, Athens, GA", "1084 Prince Ave., Athens, GA")

# Set 700 E. Broad as the starting point
l <- length(store1) # number of locations is length of vector

# set up distance as symmetric matrix
distMat1 <- matrix(0,l,l) # initialize all cells to zero
dim(distMat1) 

# compute distances and enter into cells
for(i in 1:(l-1)) { 
  for(j in (i+1):l) { 
    t <- google_distance(store1[i], store1[j], mode = c('driving'), simplify = TRUE, key = k)
    tx <- t$rows$elements[[1]]$distance$value*conv_unit(1,'m','mi') # convert meters to miles
    distMat1[i,j] <- tx
    distMat1[j,i] <- tx
    }
}
distMat1

#solving the problem store1
tsp1 <- TSP(distMat1,store1)
tour1 <- solve_TSP(tsp1,method = "nearest_insertion", start = 1)
len1 <- as.integer(tour_length(tour1))
len1
labels(tour1)


# TESTING STORE 2
store2 <- c("1650 S. Lumpkin, Athens, GA", "15 Ruthwood Lane, Athens, GA", "100 Kentucky Circle, Athens, GA", "558 West Broad St, Athens, GA", "125 Greek Park Circle, Athens, GA", "1084 Prince Ave., Athens, GA")

# Set 700 E. Broad as the starting point
l <- length(store2) # number of locations is length of vector

# set up distance as symmetric matrix
distMat2 <- matrix(0,l,l) # initialize all cells to zero
dim(distMat2) 

# compute distances and enter into cells
for(i in 1:(l-1)) { 
  for(j in (i+1):l) { 
    t <- google_distance(store2[i], store2[j], mode = c('driving'), simplify = TRUE, key = k)
    tx <- t$rows$elements[[1]]$distance$value*conv_unit(1,'m','mi') # convert meters to miles
    distMat2[i,j] <- tx
    distMat2[j,i] <- tx
    }
}
distMat2

#solving the problem store2
tsp2 <- TSP(distMat2,store2)
tour2 <- solve_TSP(tsp2,method = "nearest_insertion", start = 1)
len2 <- as.integer(tour_length(tour2))
len2
labels(tour2)


# TESTING STORE 3
store3 <- c("1800 Epps Bridge Parkway, Athens, GA", "15 Ruthwood Lane, Athens, GA", "100 Kentucky Circle, Athens, GA", "558 West Broad St, Athens, GA", "125 Greek Park Circle, Athens, GA", "1084 Prince Ave., Athens, GA")

# Set 700 E. Broad as the starting point
l <- length(store3) # number of locations is length of vector

# set up distance as symmetric matrix
distMat3 <- matrix(0,l,l) # initialize all cells to zero
dim(distMat3) 

# compute distances and enter into cells
for(i in 1:(l-1)) { 
  for(j in (i+1):l) { 
    t <- google_distance(store3[i], store3[j], mode = c('driving'), simplify = TRUE, key = k)
    tx <- t$rows$elements[[1]]$distance$value*conv_unit(1,'m','mi') # convert meters to miles
    distMat3[i,j] <- tx
    distMat3[j,i] <- tx
    }
}
distMat3

#solving the problem store3
tsp3 <- TSP(distMat3,store3)
tour3 <- solve_TSP(tsp3,method = "nearest_insertion", start = 1)
len3 <- as.integer(tour_length(tour3))
len3
labels(tour3)
```


__Which store should assigned the making of the five orders to minimize delivery distance?__

_Answer:_ 700 E. Broad, Athens, GA

__What is the shortest route in miles?__

_Answer:_ 14 miles